The Structure of Human Language Reflects Limitations of Working Memory

### 2019-04-22 paper-short.tex
some prose changes

\begin{abstract}
  %  the Nature template:
  % 1-2 sentences basic introduction, for anyone
  Each of the \~6000 natural languages currently spoken in the world embodies a different system for encoding hierarchically-structured thoughts into linear strings of words.
  By characterizing the commonalities and differences among these systems, we can gain insight into the cognitive machinery that underpins this fundamental aspect of human intelligence. 
  % 2-3 sentences detailed background, for scientists in related disciplines (linguistics, cogsci, computer science, info theory)
  In linguistics and cognitive science, there have been two

  % 1 sentence general problem
  % 1 sentence here we show
  Here we show that the order of words across human languages is structured to enable efficient transfer of information under the constraint that language must be comprehended and produced by agents with limited short-term memory capacities.
  % 2-3 sentences how main result adds to knowledge
  Using concepts from information theory, we prove a theorem showing that any efficient communication system operating under memory constraints will exhibit a property called information locality: symbols which predict each other should be produced close to each other in time.
  We show that information locality subsumes several principles of word order previously proposed in the linguistics and cognitive science literature, and explains a number of the previously-documented universal patterns in word order across languages.    
  Next, we directly show that languages are optimized to minimize memory usage by calculating lower bounds on the amount of memory resources required for language comprehension in TODO real languages and in random baseline languages, finding that the real languages require less memory.
	% TODO i'm wondering whether we should first mention the big result, and then say, well it also subsumes previously proposed constraints.
  % 1-2 sentences general context

@@

Human languages express hierarchically-structured thoughts into sentences, i.e. linear strings of words. \jd{too complicated for the general audience}
Languages differ considerably in the rules they apply to order information: English orders the object after the verb, Japanese places it before the verb.
Explaining the variation in the ways languages use to order underlying hierarchical structures into linear strings of words has been one of the foci of linguistic research (CITE).
We show that the ways in which languages order information reflect an underlying optimization principle: human languages are adapted to limitations in human working memory.

>>>>

\begin{abstract}
  %  the Nature template:
  % 1-2 sentences basic introduction, for anyone
  Each of the \~6000 natural languages currently spoken in the world embodies a different system for encoding thoughts into linear strings of words.
  By characterizing the commonalities and differences among these systems, we can gain insight into the cognitive machinery that underpins this fundamental aspect of human intelligence. 
  % 2-3 sentences detailed background, for scientists in related disciplines (linguistics, cogsci, computer science, info theory)
  In linguistics and cognitive science, there have been two

  % 1 sentence general problem
  % 1 sentence here we show
  Here we show that the order of words across human languages is structured to enable efficient transfer of information under the constraint that language must be comprehended and produced by agents with limited short-term memory capacities.
  We directly show that languages are optimized to minimize memory usage by calculating lower bounds on the amount of memory resources required for language comprehension in TODO real languages and in random baseline languages, finding that the real languages require less memory.
  % 2-3 sentences how main result adds to knowledge
  Using concepts from information theory, we prove a theorem showing that any efficient communication system operating under memory constraints will exhibit a property called information locality: symbols which predict each other should be produced close to each other in time.
  We show that information locality subsumes several principles of word order previously proposed in the linguistics and cognitive science literature, and explains a number of the previously-documented universal patterns in word order across languages.    
	% TODO i'm wondering whether we should first mention the big result, and then say, well it also subsumes previously proposed constraints.
  % 1-2 sentences general context

@@

Human languages express thoughts into sentences, i.e. linear strings of words. \jd{too complicated for the general audience}
For example, if your mountain climber friend Bob decided to turn back near the summit of Mt. Everest, you might say \emph{Bob gave the attempt to climb Mt. Everest up}.
% Run with this example

Each language embodies a different strategy for encoding thoughts into strings of words.

Languages differ considerably in the rules they apply to order information: English orders the object after the verb, Japanese places it before the verb.
Explaining the variation in the ways languages use to order underlying hierarchical structures into linear strings of words has been one of the foci of linguistic research (CITE).
We show that the ways in which languages order information reflect an underlying optimization principle: human languages are adapted to limitations in human working memory.
